---
title: "ATTEMPT Urine SOMA Scan Analysis"
author: "Ye Ji Choi and Laura Pyle"
date: "`r lubridate::today()`"
format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    code-fold: true
    embed-resources: true
    message: false
    warning: false
---

```{r echo = F, include = F}
# load required packages
library(jsonlite)
library(aws.s3)
library(dplyr)
library(ggplot2)
library(knitr)
library(ensr)
library(kableExtra)
library(ggrepel)
library(purrr)
library(tidyverse)
library(limma)
library(arsenal)
library(fgsea)
library(correlation)
library(psych)
library(openxlsx)
library(SomaPlotr)
library(lme4)
library(lmerTest)
library(here)

# source shared ATTEMPT analysis functions (from this repository)
source(here("attempt_functions.R"))

# NOTE: correlation_function.R is available in the main CHCO-Code repository
# source("correlation_function.R")

# set paths for data inputs and results outputs
# users should set these environment variables or edit paths to match local setup
bg_path <- file.path(root_path, "GSEA")
load(file.path(root_path, "KNIGHT/Somalogic data/analytes.Rdata"))

# load urine creatinine sample manifest
urine_creat <- read.xlsx(file.path(root_path, 'ATTEMPT/Data Raw/Urine Proteomics Sample Manifest Jun 25 2025.xlsx'), sheet = "Working")
names(urine_creat) <- c("position", "record_id", "visit", "box", "creatinine_u_proteomics", "comments")
# load SomaScan analyte annotation metadata
load(file.path(root_path, "Data Harmonization/Combined SomaScan/analytes_2.Rdata"))

```

```{r include = F}
# configure S3 connection to Kopah (UW research storage)
# NOTE: users must provide their own keys.json with AWS credentials
# see README for details on data access
keys <- fromJSON(Sys.getenv("ATTEMPT_KEYS_PATH", unset = "keys.json"))

Sys.setenv(
  "AWS_ACCESS_KEY_ID" = keys$MY_ACCESS_KEY,
  "AWS_SECRET_ACCESS_KEY" = keys$MY_SECRET_KEY,
  "AWS_DEFAULT_REGION" = "",
  "AWS_REGION" = "",
  "AWS_S3_ENDPOINT" = "s3.kopah.uw.edu"
)
```


```{r echo = F}
# load harmonized SomaScan dataset and filter to ATTEMPT study
harm_dat_raw <- read.csv(file.path(root_path, "Data Harmonization/Data Clean/soma_harmonized_dataset.csv"), na.strings = "")

attempt_redcap <- harm_dat_raw %>% filter(study == "ATTEMPT") %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, visit)) %>%
  dplyr::mutate(race_ethnicity_condensed = case_when(race == "White" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino White",
                                                     race == "Black or African American" &
                                                       ethnicity == "Not Hispanic or Latino" ~ "Not Hispanic or Latino Black",
                                                     ethnicity == "Hispanic or Latino" ~ "Hispanic or Latino",
                                                     T ~ "Not Hispanic or Latino Other")) %>%
  arrange(record_id) %>%
  dplyr::select(-hba1c, -starts_with("seq."))

# subset Denver site (record_id starts with 3) and recode visits to numeric weeks
attempt_redcap_denver <- attempt_redcap %>%
  filter(grepl("^3", record_id)) %>%
  mutate(visit = case_when(visit == "baseline" ~ 0,
                           visit == "4_weeks_post" ~ 4,
                           visit == "4_months_post" ~ 16,
                           visit == "screening"  ~ -4,
                           visit == "follow_up" ~ 18),
         record_id = as.integer(record_id)) %>%
  dplyr::select(-c(ethnicity, bmi, gfr_raw_plasma, gfr_bsa_plasma, date, sex, mri_date, PWV,
                   creatinine_s, cystatin_c_s, temp, pulse, height, avg_c_r2, avg_k_r2, 
                   sbp, dbp, hct, age, weight)) %>%
  rowwise() %>%
  mutate(avg_c_t1 = mean(c(t1_r_cortex, t1_l_cortex), na.rm = T),
         avg_k_t1 = mean(c(t1_r_kidney, t1_l_kidney), na.rm = T),
         tkv = (volume_left + volume_right)) %>%
  ungroup() %>%
  dplyr::rename(subject_id = record_id)

# clinical dataset from Antoine (merged version), recode visits to numeric weeks
attempt_dat <- s3readRDS("Clinical Data/ATTEMPT_AC.RDS", "attempt", region = "")
attempt_dat[attempt_dat == ""] <- NA
attempt_dat <- attempt_dat %>%
  filter(visit %in% c("baseline", "4_weeks_post", "4_months_post", "screening", "follow_up")) %>%
  mutate(visit = case_when(visit == "baseline" ~ 0,
                           visit == "4_weeks_post" ~ 4,
                           visit == "4_months_post" ~ 16,
                           visit == "screening"  ~ -4,
                           visit == "follow_up" ~ 18),
         record_id = as.integer(record_id)) %>%
  dplyr::rename(subject_id = record_id,
                treatment = treatment_arm)

# extract IDs with biopsy
attempt_clean_so_metadata <- s3readRDS("cleaned_data/attempt_clean_so_metadata.rds", "attempt", region = "")
attempt_biopsy_yn <- attempt_clean_so_metadata %>%
  dplyr::select(subject_id, visit) %>%
  distinct(subject_id, visit, .keep_all = T) %>%
  mutate(visit = case_when(visit == "PRE" ~ 0,
                           visit == "POST" ~ 16),
         biopsy_yn = "Yes")
attempt_biopsy_yn <- attempt_biopsy_yn %>%
  rbind(attempt_biopsy_yn %>%
          filter(visit == 0) %>%
          mutate(visit = -4))

# merge REDCap version (for Denver) and dataset from Antoine, add biopsy flag
attempt_dat <- left_join(attempt_dat, attempt_redcap_denver) %>%
  left_join(attempt_biopsy_yn) %>%
  mutate(biopsy_yn = case_when(biopsy_yn == "Yes" ~ "Yes",
                               T ~ "No"))
attempt_dat$treatment <- factor(attempt_dat$treatment, levels = c("Placebo", "Dapagliflozin 5mg"))
```

```{r echo = F}
# collapse SOMA data to one row per subject-visit, keep only baseline and 4-month
harm_dat <- harm_dat_raw %>%
  filter(study == "ATTEMPT") %>%
  dplyr::summarise(across(where(negate(is.numeric)), ~ ifelse(all(is.na(.x)), NA_character_, last(na.omit(.x)))),
                   across(where(is.numeric), ~ ifelse(all(is.na(.x)), NA_real_, mean(.x, na.rm = TRUE))),
                   .by = c(record_id, visit)) %>%
  mutate(visit = case_when(visit == "baseline" ~ 0,
                           visit == "4_weeks_post" ~ 4,
                           visit == "4_months_post" ~ 16,
                           visit == "screening"  ~ -4,
                           visit == "follow_up" ~ 18),
         record_id_visit = paste0(record_id, "_", visit),
         treatment_arm = case_when(treatment_arm == "Placebo" ~ "Placebo",
                                   T ~ "Dapagliflozin")) %>%
  filter(visit %in% c(0, 16))

# filter to samples with urine SOMA data, exclude very low creatinine (likely serum mix-up)
soma <- harm_dat %>%
  filter(!is.na(seq.10030.8_urine)) %>%
  filter(urine_creat_proteomics > 1)

soma$visit <- factor(soma$visit, levels = c(0, 16), label = c("PRE", "POST"))
```

# SOMA

```{r echo = F, results='asis'}
# demographics table stratified by treatment arm and visit
summary(arsenal::tableby(visit ~ age + sex + bmi + diabetes_duration,
                         data = soma, strata = treatment_arm), test = F,
        labelTranslations = c("age" = "Age", "sex" = "Sex", "acr_u" = "UACR",
                              "bmi" = "BMI", "diabetes_duration" = "Diabetes duration"))
```

## Placebo
```{r echo = F}
# fit limma within-treatment model for placebo (PRE vs POST)
placebo_res_urine <- run_limma_proteomics(data = soma,
                                          analyte_info = analytes_attempt,
                                          treatment = "Placebo",
                                          feature_suffix = "_urine_cradj",
                                          model_type = "within_treatment")
```

```{r echo = F}
# volcano plots for placebo within-treatment results (raw and adj p-values)
plot_volcano_proteomics(data = placebo_res_urine$results_annotated, fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive after placebo treatment",
                        negative_text = "Negative after placebo treatment",
                        formula = "visit",
                        file_suffix = "_placebo_urine_proteomics")
plot_volcano_proteomics(data = placebo_res_urine$results_annotated, fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive after placebo treatment",
                        negative_text = "Negative after placebo treatment",
                        formula = "visit",
                        file_suffix = "_placebo_urine_proteomics_adjp")
```

## Dapagliflozin
```{r echo = F}
# fit limma within-treatment model for dapagliflozin (PRE vs POST)
dapa_res_urine <- run_limma_proteomics(data = soma,
                                       analyte_info = analytes_attempt,
                                       treatment = "Dapagliflozin",
                                       feature_suffix = "_urine_cradj",
                                       model_type = "within_treatment")
```

```{r echo = F}
# volcano plots for dapagliflozin within-treatment results (raw and adj p-values)
plot_volcano_proteomics(data = dapa_res_urine$results_annotated, fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive after dapagliflozin treatment",
                        negative_text = "Negative after dapagliflozin treatment",
                        formula = "visit",
                        file_suffix = "_dapa_urine_proteomics")
plot_volcano_proteomics(data = dapa_res_urine$results_annotated, fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive after dapagliflozin treatment",
                        negative_text = "Negative after dapagliflozin treatment",
                        formula = "visit",
                        file_suffix = "_dapa_urine_proteomics_adjp")
```

## PRE (Placebo vs. Dapa)
```{r echo = F}
# fit limma within-visit model at baseline (placebo vs dapa)
pre_res_urine <- run_limma_proteomics(data = soma,
                                      analyte_info = analytes_attempt,
                                      visit = "PRE",
                                      feature_suffix = "_urine_cradj",
                                      model_type = "within_visit")
```

```{r echo = F}
# volcano plots for baseline treatment comparison (raw and adj p-values)
plot_volcano_proteomics(data = pre_res_urine$results_annotated, fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive in dapagliflozin\ncompared to placebo (at baseline)",
                        negative_text = "Negative in dapagliflozin\ncompared to placebo (at baseline)",
                        formula = "treatment",
                        file_suffix = "_pre_urine_proteomics")
plot_volcano_proteomics(data = pre_res_urine$results_annotated, fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive in dapagliflozin\ncompared to placebo (at baseline)",
                        negative_text = "Negative in dapagliflozin\ncompared to placebo (at baseline)",
                        formula = "treatment",
                        file_suffix = "_pre_urine_proteomics_adjp")
```

## POST (Placebo vs. Dapa)
```{r echo = F}
# fit limma within-visit model at follow-up (placebo vs dapa)
post_res_urine <- run_limma_proteomics(data = soma,
                                       analyte_info = analytes_attempt,
                                       visit = "POST",
                                       feature_suffix = "_urine_cradj",
                                       model_type = "within_visit")
```

```{r echo = F}
# volcano plots for follow-up treatment comparison (raw and adj p-values)
plot_volcano_proteomics(data = post_res_urine$results_annotated, fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC",
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive in dapagliflozin\ncompared to placebo (at follow-up)",
                        negative_text = "Negative in dapagliflozin\ncompared to placebo (at follow-up)",
                        formula = "treatment",
                        file_suffix = "_post_urine_proteomics")
plot_volcano_proteomics(data = post_res_urine$results_annotated, fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive in dapagliflozin\ncompared to placebo (at follow-up)",
                        negative_text = "Negative in dapagliflozin\ncompared to placebo (at follow-up)",
                        formula = "treatment",
                        file_suffix = "_post_urine_proteomics_adjp")
```

# DiD (Normalized by creatinine)

```{r echo = F, eval = F}
# using creatinine adjusted proteins
did_res_urine <- run_limma_proteomics(data = soma,
                                      analyte_info = analytes_attempt,
                                      feature_suffix = "_urine_cradj",
                                      model_type = "interaction")

# fit DiD mixed model with random intercept for subject
did_res_urine_mixed <- run_limma_proteomics(data = soma,
                                            analyte_info = analytes_attempt,
                                            feature_suffix = "_urine_cradj",
                                            model_type = "interaction_random")
# 6 out of 7289 singularity warning (0.1%)

s3saveRDS(did_res_urine_mixed, "Proteomics/urine_cradj_proteomics_res.RDS", "attempt", region = "")
write.csv(did_res_urine_mixed$results_annotated, file.path(results_dir, "Somascan/urine_cradj_proteomics_res.csv"), row.names = F)
```

```{r echo = F}
# load saved DiD mixed model results and create volcano plots
did_res_urine_mixed <- s3readRDS("Proteomics/urine_cradj_proteomics_res.RDS", "attempt", region = "")

plot_volcano_proteomics(data = did_res_urine_mixed$results_annotated, 
                        fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive with Dapagliflozin",
                        negative_text = "Negative with Dapagliflozin",
                        formula = "treatment:visit + (1|subject)",
                        file_suffix = "_DiD_urine_proteomics_mixed")

urine_volcano <- plot_volcano_proteomics(data = did_res_urine_mixed$results_annotated, 
                        fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(FDR adj.p-value)",
                        positive_text = "Positive with Dapagliflozin",
                        negative_text = "Negative with Dapagliflozin",
                        formula = "treatment*visit + (1|subject)",
                        file_suffix = "_DiD_urine_proteomics_adjp_mixed")
```

```{r echo = F}
# pivot DiD results wide for heatmap, filter to FDR < 0.05
res_full_wide <- did_res_urine_mixed$results_full %>%
  dplyr::select(-effect) %>%
  # Pivot wider, creating columns for each term's statistics
  tidyr::pivot_wider(
    id_cols = c(AptName, SeqId, EntrezGeneSymbol, Target),  # Keep these as row identifiers
    names_from = term,
    values_from = c(logFC, std.error, statistic, df, P.Value, adj.P.Val),
    names_glue = "{term}_{.value}"  # Format: term_statistic
  ) %>%
  mutate(EntrezGeneSymbol = case_when(EntrezGeneSymbol == "" ~ Target, T ~ EntrezGeneSymbol),
         Gene_seq = paste0(EntrezGeneSymbol, " (", SeqId, ")")) %>%
  filter(`treatment_armDapagliflozin:visitPOST_adj.P.Val` < 0.05)

# heatmap of top 20 significant DiD proteins
urine_heatmap <- plot_treatment_heatmap(data = res_full_wide,
                       p_col = "P.Value",
                       top_n = 20,
                       logfc_col = "treatment_armDapagliflozin:visitPOST_logFC",
                       fdr_col = "treatment_armDapagliflozin:visitPOST_adj.P.Val",
                       logfc_visit = "visitPOST_logFC",
                       gene_col = "Gene_seq",
                       heatmap_caption = F)

# combine volcano and heatmap into panel figure
ggarrange(urine_volcano, urine_heatmap, widths = c(2,1), labels = "AUTO", hjust = c(0,0.5))
ggsave(file.path(results_dir, "Figures/Panels/urine_proteomics_panel.png"), width = 10, height = 6)

```

## GSEA (fgsea)

```{r echo = F}
# run GSEA pathway analysis on creatinine-adjusted DiD results
soma_DiD_fgsea <- run_fgsea_analysis(bg_path = bg_path,
                                     results_annotated = did_res_urine_mixed$results_annotated,
                                     stat_col = "statistic")
soma_DiD_fgsea$summary
```

```{r echo = F}
# plot top 30 enriched pathways for KEGG, REACTOME, and GO
# KEGG
plot_fgsea_transpose(soma_DiD_fgsea$kegg,
                     title = "SOMA DiD Top 30 KEGG",
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_SOMA_DiD_KEGG.png"),
       width = 27.5, height = 14, scale = 1)
# REACTOME
plot_fgsea_transpose(soma_DiD_fgsea$reactome,
                     title = "SOMA DiD Top 30 REACTOME",
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_SOMA_DiD_REACTOME.png"),
       width = 27.5, height = 14, scale = 1)
# GO
plot_fgsea_transpose(soma_DiD_fgsea$go,
                     title = "SOMA DiD Top 30 GO",
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_SOMA_DiD_GO.png"),
       width = 27.5, height = 14, scale = 1)
```

## Integrate transcriptomics

```{r echo = F}
# looking for direct protein shedding/secretion into urine
# read in results from transcriptomics (PT, TAL, DCT, CD (collecting duct = PC, IC), and POD for urine proteimics)
celltypes <- c("PT", "TAL", "PC", "IC", "POD")

# load and process nebula DiD results for each cell type (full gene set)
for (cell in celltypes)  {
  input_path <- paste0(cell,"/nebula/",tolower(cell),"_attempt_nebula_res_reml_pooled.rds")
  cell_df <- s3readRDS(input_path, bucket = "attempt", region = "")
  processed <- process_nebula_results(cell_df)
  lower_cell <- tolower(cell)
  var_name <- paste0(lower_cell, "_nebula_res_combined_reml_pooled")
  
  assign(var_name, processed$results)
}

# filter to nominally significant DiD proteins
urine_did_sig <- did_res_urine_mixed$results_annotated %>%
  filter(P.Value < 0.05)

# compute PRE-to-POST delta for each protein per subject
urine_delta <- soma %>%
  dplyr::select(record_id, visit, treatment_arm, starts_with("seq") & ends_with("cradj")) %>%
  pivot_wider(
    names_from = visit,
    values_from = starts_with("seq"),
    names_glue = "{.value}_{visit}"
  ) %>%
  mutate(across(ends_with("_POST"), 
                ~ . - get(sub("_POST$", "_PRE", cur_column())),
                .names = "delta_{.col}")) %>%
  dplyr::select(record_id, treatment_arm, starts_with("delta_"))

s3saveRDS(urine_delta, "Proteomics/urine_cradj_proteomics_delta.RDS", "attempt", region = "")
```

```{r echo = F}
# map overlapping significant genes between scRNA DiD and urine protein DiD per cell type
walk(celltypes, function(ct) {
  
  # Dynamically get the nebula results object based on cell type
  nebula_obj <- get(paste0(tolower(ct), "_nebula_res_combined_reml_pooled"))
  
  sig <- nebula_obj %>%
    filter(`p_treatmentDapagliflozin:visitPOST` < 0.05)
  
  map <- analytes_attempt %>%
    filter(EntrezGeneSymbol %in% unique(sig$Gene[sig$Gene %in% urine_did_sig$EntrezGeneSymbol])) %>%
    dplyr::select(Target, TargetFullName, AptName, EntrezGeneSymbol)
  
  write.csv(map, 
            file.path(results_dir, "Somascan", paste0(ct, "_urine_proteomics_map.csv")),
            row.names = FALSE, na = "")
})
```


## Correlations with clinical variables

### mGFR

```{r echo = F, warning=FALSE, quiet = TRUE}
# correlate delta urine proteins with delta mGFR (labeled by target name)
urine_soma_mgfr <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "mgfr_jodal",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 mGFR",
  pre_visit = -4,
  label = "Target"
)

# correlation table
urine_soma_mgfr$correlation_table

# plots
urine_soma_mgfr$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_mgfr.png"),
       urine_soma_mgfr$plot_all, width = 10, height = 7)

# same mGFR correlation but labeled by gene symbol
urine_soma_mgfr_gene <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "mgfr_jodal",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 mGFR",
  pre_visit = -4,
  label = "EntrezGeneSymbol"
)

# correlation table
urine_soma_mgfr_gene$correlation_table

# plots
urine_soma_mgfr_gene$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_mgfr_gene.png"),
       urine_soma_mgfr_gene$plot_all, width = 10, height = 7)
```

### mGFR adjusted

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta BSA-adjusted mGFR
urine_soma_mgfr_adj <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "mgfr_jodal_bsa",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 mGFR (BSA)",
  pre_visit = -4
)

# correlation table
urine_soma_mgfr_adj$correlation_table

# plots
urine_soma_mgfr_adj$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_mgfr_bsa.png"),
       urine_soma_mgfr_adj$plot_all, width = 10, height = 7)
```

### HbA1c

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta HbA1c
urine_soma_hba1c <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "hba1c",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 HbA1c"
)

# correlation table
urine_soma_hba1c$correlation_table

# plots
urine_soma_hba1c$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_hba1c.png"),
       urine_soma_hba1c$plot_all, width = 10, height = 7)
```

### TIR

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta time in range
urine_soma_tir <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "cgm_tir",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 TIR"
)

# correlation table
urine_soma_tir$correlation_table

# plots
urine_soma_tir$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_tir.png"),
       urine_soma_tir$plot_all, width = 10, height = 7)
```

### Weight

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta weight
urine_soma_weight <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "weight",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Weight"
)

# correlation table
urine_soma_weight$correlation_table

# plots
urine_soma_weight$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_weight.png"),
       urine_soma_weight$plot_all, width = 10, height = 7)
```

### PWV

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta pulse wave velocity
urine_soma_PWV <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "PWV",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 PWV",
  pre_visit = -4
)

# correlation table
urine_soma_PWV$correlation_table

# plots
urine_soma_PWV$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_PWV.png"),
       urine_soma_PWV$plot_all, width = 10, height = 7)
```

### Cortical R2*

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta cortical R2* (BOLD MRI oxygenation)
urine_soma_avg_c_r2 <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "avg_c_r2",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Cortical R2*",
  pre_visit = -4
)

# correlation table
urine_soma_avg_c_r2$correlation_table

# plots
urine_soma_avg_c_r2$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_avg_c_r2.png"),
       urine_soma_avg_c_r2$plot_all, width = 10, height = 7)
```

### Kidney R2*

```{r echo = F, warning=FALSE}
# correlate delta urine proteins with delta kidney R2* (BOLD MRI oxygenation)
urine_soma_avg_k_r2 <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "avg_k_r2",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Kidney R2*",
  pre_visit = -4
)

# correlation table
urine_soma_avg_k_r2$correlation_table

# plots
urine_soma_avg_k_r2$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_spearman_urine_avg_k_r2.png"),
       urine_soma_avg_k_r2$plot_all, width = 10, height = 7)
```

# DiD (Creatinine in model)

```{r echo = F}
# fit DiD fixed-effect model with creatinine as covariate (not pre-normalized)
did_res_urine_cr <- run_limma_proteomics(data = soma,
                                         analyte_info = analytes_attempt,
                                         feature_suffix = "_urine",
                                         creatinine = "urine_creat_proteomics",
                                         model_type = "interaction")

# fit DiD mixed model with creatinine as covariate and random subject intercept
did_res_urine_mixed_cr <- run_limma_proteomics(data = soma,
                                               analyte_info = analytes_attempt,
                                               feature_suffix = "_urine",
                                               creatinine = "urine_creat_proteomics",
                                               model_type = "interaction_random")
# 316 out of 7289 singularity warning (4.3%)
```

```{r echo = F}
# volcano plots for DiD with creatinine-in-model (fixed and mixed, raw and adj p)
plot_volcano_proteomics(data = did_res_urine_cr$results_annotated,
                        fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC",
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive DiD",
                        negative_text = "Negative DiD",
                        formula = "treatment:visit",
                        file_suffix = "_DiD_urine_proteomics_creatinine")

plot_volcano_proteomics(data = did_res_urine_mixed_cr$results_annotated, 
                        fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive DiD",
                        negative_text = "Negative DiD",
                        formula = "treatment:visit",
                        file_suffix = "_DiD_urine_proteomics_adjp_creatinine")

plot_volcano_proteomics(data = did_res_urine_mixed_cr$results_annotated, 
                        fc = "logFC",
                        p_col = "P.Value",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(p-value)",
                        positive_text = "Positive DiD",
                        negative_text = "Negative DiD",
                        formula = "treatment:visit + (1|subject)",
                        file_suffix = "_DiD_urine_proteomics_mixed_creatinine")

plot_volcano_proteomics(data = did_res_urine_mixed_cr$results_annotated, 
                        fc = "logFC",
                        p_col = "adj.P.Val",
                        title = NULL,
                        x_axis = "LogFC", 
                        y_axis = "-log10(adj.p-value)",
                        positive_text = "Positive DiD",
                        negative_text = "Negative DiD",
                        formula = "treatment:visit + (1|subject)",
                        file_suffix = "_DiD_urine_proteomics_adjp_mixed_creatinine")
```

## GSEA (fgsea)

```{r echo = F}
# run GSEA pathway analysis on creatinine-in-model DiD results
soma_DiD_fgsea <- run_fgsea_analysis(bg_path = bg_path,
                                     results_annotated = did_res_urine_mixed_cr$results_annotated,
                                     stat_col = "statistic")
soma_DiD_fgsea$summary
```

```{r echo = F}
# plot top 30 enriched pathways for KEGG, REACTOME, and GO (creatinine-in-model)
# KEGG
plot_fgsea_transpose(soma_DiD_fgsea$kegg,
                     title = "SOMA DiD Top 30 KEGG",
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_creatinine_SOMA_DiD_KEGG.png"),
       width = 27.5, height = 14, scale = 1)
# REACTOME
plot_fgsea_transpose(soma_DiD_fgsea$reactome, 
                     title = "SOMA DiD Top 30 REACTOME", 
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_creatinine_SOMA_DiD_REACTOME.png"),
       width = 27.5, height = 14, scale = 1)
# GO
plot_fgsea_transpose(soma_DiD_fgsea$go, 
                     title = "SOMA DiD Top 30 GO", 
                     xmax = 30)
ggsave(file.path(results_dir, "Figures/Proteomics/Pathways/urine_creatinine_SOMA_DiD_GO.png"),
       width = 27.5, height = 14, scale = 1)
```

## Correlations with clinical variables

### mGFR

```{r echo = F, warning=FALSE, quiet = TRUE}
# correlate creatinine-in-model DiD proteins with delta mGFR
urine_soma_mgfr <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "mgfr_jodal",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 mGFR",
  pre_visit = -4
)

# correlation table
urine_soma_mgfr$correlation_table

# plots
urine_soma_mgfr$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_mgfr.png"),
       urine_soma_mgfr$plot_all, width = 10, height = 7)
```

### mGFR adjusted

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta BSA-adjusted mGFR
urine_soma_mgfr_adj <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "mgfr_jodal_bsa",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 mGFR (BSA)",
  pre_visit = -4
)

# correlation table
urine_soma_mgfr_adj$correlation_table

# plots
urine_soma_mgfr_adj$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_mgfr_bsa.png"),
       urine_soma_mgfr_adj$plot_all, width = 10, height = 7)
```

### HbA1c

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta HbA1c
urine_soma_hba1c <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "hba1c",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 HbA1c"
)

# correlation table
urine_soma_hba1c$correlation_table

# plots
urine_soma_hba1c$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_hba1c.png"),
       urine_soma_hba1c$plot_all, width = 10, height = 7)
```
### TIR

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta time in range
urine_soma_tir <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "cgm_tir",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 TIR"
)

# correlation table
urine_soma_tir$correlation_table

# plots
urine_soma_tir$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_tir.png"),
       urine_soma_tir$plot_all, width = 10, height = 7)
```

### Weight

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta weight
urine_soma_weight <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "weight",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Weight"
)

# correlation table
urine_soma_weight$correlation_table

# plots
urine_soma_weight$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_weight.png"),
       urine_soma_weight$plot_all, width = 10, height = 7)
```

### PWV

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta PWV
urine_soma_PWV <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "PWV",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 PWV",
  pre_visit = -4
)

# correlation table
urine_soma_PWV$correlation_table

# plots
urine_soma_PWV$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_PWV.png"),
       urine_soma_PWV$plot_all, width = 10, height = 7)
```

### Cortical R2*

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta cortical R2*
urine_soma_avg_c_r2 <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "avg_c_r2",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Cortical R2*",
  pre_visit = -4
)

# correlation table
urine_soma_avg_c_r2$correlation_table

# plots
urine_soma_avg_c_r2$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_avg_c_r2.png"),
       urine_soma_avg_c_r2$plot_all, width = 10, height = 7)
```

### Kidney R2*

```{r echo = F, warning=FALSE}
# correlate creatinine-in-model DiD proteins with delta kidney R2*
urine_soma_avg_k_r2 <- soma_corr(
  clinical_raw = attempt_dat,
  soma_raw     = soma,
  did_df       = did_res_urine_mixed_cr$results_annotated,
  analytes_info= analytes_attempt,
  clinical_var = "avg_k_r2",
  feature_suffix = "_urine_cradj",
  proper_name = "\u0394 Kidney R2*",
  pre_visit = -4
)

# correlation table
urine_soma_avg_k_r2$correlation_table

# plots
urine_soma_avg_k_r2$plot_all
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/Correlations/volcano_mixed_cr_spearman_urine_avg_k_r2.png"),
       urine_soma_avg_k_r2$plot_all, width = 10, height = 7)
```

# Integration with scRNA (results from Hyak)

```{r include = F}
# define cell type groupings for pseudobulk scRNA analysis
celltype_groups <- list(
  PT = c("PT-S1/S2", "PT-S3", "aPT"),
  TAL = c("C-TAL-1", "C-TAL-2", "aTAL", "dTAL"),
  PC = c("CCD-PC", "CNT-PC", "dCCD-PC", "M-PC", "tPC-IC"),
  EC = c("EC-AVR", "EC-GC", "EC-PTC", "EC-AEA", "EC-LYM", "EC/VSMC"),
  IC = c("IC-A", "IC-B", "aIC"),
  Immune = c("MAC", "MON", "cDC", "pDC", "CD4+ T", "CD8+ T", "B", "NK"),
  Immune_myeloid = c("MAC", "MON", "cDC", "pDC"),
  Immune_lymphoid = c("CD4+ T", "CD8+ T", "B", "NK"),
  VSMC_P_FIB = c("VSMC/P", "FIB"),
  POD = "POD",
  DCT = c("DCT", "dDCT"),  # Distal convoluted tubule
  CD = c("CCD-PC", "dCCD-PC", "CNT-PC", "tPC-IC", "MC", "aIC", "IC-A", "IC-B", "CNT")  # Collecting duct & connecting tubule
)

celltype_groups$Combined_u <- c(
  celltype_groups$PT,
  celltype_groups$TAL,
  celltype_groups$DCT,
  celltype_groups$CD,
  celltype_groups$POD
)

# read nebula mixed model results from Hyak for grouped cell types (full gene set)
for (cell in names(celltype_groups))  {
  cell <- gsub("/", "_", cell)

  annotated_df <- read.csv(file.path(results_dir, "nebula", paste0(tolower(cell), "_kpmp_res.csv")))
  colnames(annotated_df) <- gsub("\\.", ":", colnames(annotated_df))
  cell <- gsub("-", "_", cell)
  cell <- gsub("\\+\\s", "_", cell)
  lower_cell <- tolower(cell)
  var_name <- paste0(lower_cell, "_kpmp")
  assign(var_name, annotated_df)
}

# read nebula results for individual sub-celltypes (top 2000 HVG)
for (cell in unlist(celltype_groups))  {
  cell <- gsub("/", "_", cell)
  cell <- gsub("-", "_", cell)
  cell <- gsub("\\+\\s", "_", cell)

  annotated_df <- read.csv(file.path(results_dir, "nebula", paste0(tolower(cell), "_kpmp_hvg_res.csv")))
  colnames(annotated_df) <- gsub("\\.", ":", colnames(annotated_df))
  cell <- gsub("-", "_", cell)
  cell <- gsub("\\+\\s", "_", cell)
  lower_cell <- tolower(cell)
  var_name <- paste0(lower_cell, "_hvg_kpmp")
  assign(var_name, annotated_df)
}
```


```{r echo = F}
# load PT transcript-protein association results from nebula
pt_urine_raw <- s3readRDS("Proteomics/pt_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
pt_urine <- process_nebula_results(pt_urine_raw$results, pval_col = "p_log2(protein)")$results
pt_urine <- pt_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
# genes with significant transcript-protein association
pt_urine_genes <- pt_urine %>%
  filter(`p_log2(protein)` < 0.05) %>%
  pull(Gene)

# reload urine DiD results for merging
did_res_urine_mixed <- s3readRDS("Proteomics/urine_cradj_proteomics_res.RDS", "attempt", region = "")

# get urine protein DiD logFC for genes with significant transcript-protein association
did_urine_pt <- did_res_urine_mixed$results_annotated %>%
  filter(EntrezGeneSymbol %in% pt_urine_genes) %>%
  dplyr::select(Gene = EntrezGeneSymbol, SeqId, logFC_urine = logFC, p.value_urine = P.Value)

# get transcript DiD logFC for those same genes
did_trans_pt_u <- pt_kpmp %>%
  filter(Gene %in% pt_urine_genes) %>%
  dplyr::select(Gene, logFC_transcript = `logFC_treatmentDapagliflozin:visitPOST`, p.value_transcript = `p_treatmentDapagliflozin:visitPOST`)

# merge and classify concordance direction (U+/T+, U-/T-, discordant, NS)
urine_pt_merge <- full_join(did_urine_pt, did_trans_pt_u) %>%
  mutate(direction = case_when(logFC_urine > 0 & p.value_urine < 0.05 & logFC_transcript > 0 & p.value_transcript < 0.05 ~ "U+/T+",
                               logFC_urine < 0 & p.value_urine < 0.05 & logFC_transcript < 0 & p.value_transcript < 0.05 ~ "U-/T-",
                               logFC_urine > 0 & p.value_urine < 0.05 & logFC_transcript < 0 & p.value_transcript < 0.05 ~ "U+/T-",
                               logFC_urine < 0 & p.value_urine < 0.05 & logFC_transcript > 0 & p.value_transcript < 0.05 ~ "U-/T+",
                               T ~ "NS"))


theme_transparent <- theme(
  plot.background   = element_rect(fill = "transparent", color = NA),
  panel.background  = element_rect(fill = "transparent", color = NA),
  legend.background = element_rect(fill = "transparent", color = NA)
)

# label top 20 concordant/discordant genes in each direction for volcano
pos_labels <- urine_pt_merge %>%
  filter(direction != "NS" & logFC_urine > 0) %>%
  arrange(p.value_urine) %>%
  head(20)
neg_labels <- urine_pt_merge %>%
  filter(direction != "NS" & logFC_urine < 0) %>%
  arrange(p.value_urine) %>%
  head(20)
labels_dat <- rbind(pos_labels, neg_labels)

# volcano plot showing urine protein DiD colored by transcript concordance (PT)
urine_pt_merge %>%
  ggplot(aes(x = logFC_urine, y = -log10(p.value_urine), color = direction)) +
  geom_hline(yintercept = -log10(0.05), color = "darkgrey", linetype = "dashed") +
  geom_point(alpha = 0.5) +
  geom_text_repel(data = labels_dat, aes(label = Gene, color = direction), alpha = 0.8,
                    max.overlaps = Inf) +
  scale_color_manual(values = c("U+/T+" = "#588157",
                                "U-/T-" = "#588157",
                                "U+/T-" = "#457b9d",
                                "U-/T+" = "#f28482",
                                "NS" = "lightgrey")) +
  theme_transparent +
  labs(x = "logFC(Urine proteomics)", y = "-log10(P.Value)",
       color = "Urine/Transcript")
```

```{r echo = F}
# volcano for PT transcript-protein association and directional concordance plot
plot_volcano(pt_urine,
             fc = "logFC_log2(protein)", 
             p_col = "p_log2(protein)", 
             x_axis = "logFC log2(protein)", 
             y_axis = "-log10(p.value)", 
             file_suffix = "_kpmp_pt_urine",
             formula = "log2(protein) + treatment", 
             cell_type = "PT",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2)

create_directional_volcano(
  rna_data = pt_kpmp,
  protein_data = pt_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "PT"
)
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_pt.png"))
```

```{r echo = F}
# quantify PT transcript-protein concordance for urine proteomics
pt_concordance_urine <- concordance_by_celltype(
  biofluid_df = pt_urine,
  scrna_df    = pt_kpmp,
  biofluid_label = "urine",
  celltype = "PT",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

```{r echo = F}
# load and process TAL transcript-protein association results
tal_urine_raw <- s3readRDS("Proteomics/tal_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
tal_urine <- process_nebula_results(tal_urine_raw$results, pval_col = "p_log2(protein)")$results
tal_urine <- tal_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
tal_urine <- subset(tal_urine, `logFC_log2(protein)` > -1) # 2 outliers
```

```{r echo = F}
# volcano for TAL transcript-protein association and directional concordance plot
plot_volcano(tal_urine, fc = "logFC_log2(protein)", p_col = "p_log2(protein)", x_axis = "logFC log2(protein)", y_axis = "-log10(p.value)", file_suffix = "_kpmp_tal_urine",
             formula = "log2(protein) + treatment", 
             cell_type = "TAL",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2)

create_directional_volcano(
  rna_data = tal_kpmp,
  protein_data = tal_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "TAL"
)
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_tal.png"))
```

```{r echo = F}
# quantify TAL transcript-protein concordance for urine proteomics
tal_concordance_urine <- concordance_by_celltype(
  biofluid_df = tal_urine,
  scrna_df    = tal_kpmp,
  biofluid_label = "urine",
  celltype = "TAL",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

```{r echo = F}
# load and process DCT transcript-protein association results
dct_urine_raw <- s3readRDS("Proteomics/dct_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
dct_urine <- process_nebula_results(dct_urine_raw$results, pval_col = "p_log2(protein)")$results
dct_urine <- dct_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
dct_urine <- subset(dct_urine, `logFC_log2(protein)` < 1) # 2 outliers
```

```{r echo = F}
# volcano for DCT transcript-protein association and directional concordance plot
plot_volcano(dct_urine, fc = "logFC_log2(protein)", p_col = "p_log2(protein)", x_axis = "logFC log2(protein)", y_axis = "-log10(p.value)", file_suffix = "_kpmp_dct_urine",
             formula = "log2(protein) + treatment", cell_type = "DCT",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2)

create_directional_volcano(
  rna_data = dct_kpmp,
  protein_data = dct_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "DCT"
)
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_dct.png"))
```

```{r echo = F}
# quantify DCT transcript-protein concordance for urine proteomics
dct_concordance_urine <- concordance_by_celltype(
  biofluid_df = dct_urine,
  scrna_df    = dct_kpmp,
  biofluid_label = "urine",
  celltype = "DCT",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

```{r echo = F}
# load and process CD (collecting duct) transcript-protein association results
cd_urine_raw <- s3readRDS("Proteomics/cd_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
cd_urine <- process_nebula_results(cd_urine_raw$results, pval_col = "p_log2(protein)")$results
cd_urine <- cd_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
```

```{r echo = F}
# volcano for CD transcript-protein association and directional concordance plot
plot_volcano(cd_urine, fc = "logFC_log2(protein)", p_col = "p_log2(protein)", x_axis = "logFC log2(protein)", y_axis = "-log10(p.value)", file_suffix = "_kpmp_cd_urine",
             formula = "log2(protein) + treatment", cell_type = "CD",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2)

create_directional_volcano(
  rna_data = cd_kpmp,
  protein_data = cd_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "CD"
)
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_cd.png"))
```

```{r echo = F}
# quantify CD transcript-protein concordance for urine proteomics
cd_concordance_urine <- concordance_by_celltype(
  biofluid_df = cd_urine,
  scrna_df    = cd_kpmp,
  biofluid_label = "urine",
  celltype = "CD",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

```{r echo = F}
# load and process POD (podocyte) transcript-protein association results
pod_urine_raw <- s3readRDS("Proteomics/pod_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
pod_urine <- process_nebula_results(pod_urine_raw$results, pval_col = "p_log2(protein)")$results
pod_urine <- pod_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
```

```{r echo = F}
# volcano for POD transcript-protein association and directional concordance plot
plot_volcano(pod_urine, fc = "logFC_log2(protein)", p_col = "p_log2(protein)", x_axis = "logFC log2(protein)", y_axis = "-log10(p.value)", file_suffix = "_kpmp_pod_urine",
             formula = "log2(protein) + treatment", cell_type = "POD",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2)

create_directional_volcano(
  rna_data = pod_kpmp,
  protein_data = pod_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "POD"
)
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_pod.png"))
```

```{r echo = F}
# quantify POD transcript-protein concordance for urine proteomics
pod_concordance_urine <- concordance_by_celltype(
  biofluid_df = pod_urine,
  scrna_df    = pod_kpmp,
  biofluid_label = "urine",
  celltype = "POD",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

```{r echo = F}
# load and process combined (PT+TAL+DCT+CD+POD) transcript-protein results
combined_urine_raw <- s3readRDS("Proteomics/combined_nebula_urine_soma_res_kpmp.rds", "attempt", region = "")
combined_urine <- process_nebula_results(combined_urine_raw$results, pval_col = "p_log2(protein)")$results
combined_urine <- combined_urine %>%
  mutate(Gene = sub("_.*", "", Gene))
```

```{r echo = F}
# volcano for combined cell type transcript-protein association and concordance
plot_volcano(combined_urine,
             fc = "logFC_log2(protein)", 
             p_col = "p_log2(protein)", 
             x_axis = "logFC log2(protein)", 
             y_axis = "-log10(p.value)", 
             file_suffix = "_kpmp_combined_urine",
             formula = "log2(protein) + treatment", 
             cell_type = "Combined (PT, TAL, DCT, CD, POD)",
             positive_text = "Positively associated\nwith transcript",
             negative_text = "Negatively associated\nwith transcript") + 
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 0.2) +
  coord_cartesian(clip = "off")

create_directional_volcano(
  rna_data = combined_u_kpmp,
  protein_data = combined_urine %>% filter(`logFC_log2(protein)` > -10),
  cell_type_name = "Combined (PT, TAL, DCT, CD, POD)"
) + coord_cartesian(clip = "off")
ggsave(file.path(results_dir, "Figures/Proteomics/Volcano Plots/urine_combined.png"))
```

```{r echo = F}
# quantify combined transcript-protein concordance for urine proteomics
combined_concordance_urine <- concordance_by_celltype(
  biofluid_df = combined_urine,
  scrna_df    = combined_u_kpmp,
  biofluid_label = "urine",
  celltype = "Combined",
  out_dir = file.path(results_dir, "Somascan/Transcript concordance")
)
```

# Looking for overlap between treatment DiD & proteomics DiD

```{r echo = F}
# redefine cell type groups for treatment DiD overlap analysis (PT, IC, POD only)
celltype_groups <- list(
  PT = c("PT-S1/S2", "PT-S3", "aPT"),
  # TAL = c("C-TAL-1", "C-TAL-2", "aTAL", "dTAL"),
  # PC = c("CCD-PC", "CNT-PC", "dCCD-PC", "M-PC", "tPC-IC"),
  IC = c("IC-A", "IC-B", "aIC"),
  # DCT = c("DCT","dDCT"),
  POD = "POD"
)

# load grouped nebula treatment DiD results for PT, IC, POD
for (cell in names(celltype_groups))  {
  input_path <- paste0("grouped/", cell, "/ALL/nebula/", "grouped_", tolower(cell), "_kpmp_attempt_nebula_res_reml_pooled.rds")
  cell_df <- s3readRDS(input_path, bucket = "attempt", region = "")

  processed <- process_nebula_results(cell_df)
  lower_cell <- tolower(cell)
  var_name <- paste0(lower_cell, "_kpmp")

  assign(var_name, processed$results)
}
```

```{r echo = F}
# batch concordance analysis: overlap between urine protein DiD and transcript DiD
urine_res <- batch_concordance_analysis(
  protein_data = did_res_urine_mixed$results_annotated,  # This is correct
  transcript_data_list = list(PT = pt_kpmp, IC = ic_kpmp, POD = pod_kpmp),
  protein_source = 'urine',
  output_dir = file.path(results_dir, "Figures/Pathways/enrichr")
)
```